\chapwithtoc{Conclusion}

In this thesis, we have proposed a design of an experiment offering the possibility of assembling a stable Coulomb crystal with electrons trapped inside it. We are aiming to achieve cooling of electrons up to the point when they form a Fermi gas. If successful, such an experiment would expand current lines of research on quantum systems. 

We have presented a comprehensive overview of the current state of knowledge regarding the trapping of charged particles solely by an electric field. We have shown the general treatment of effective potential and stability, using it on the concrete case of the ideal Paul trap, for which we have derived equations of motion. Based on pseudopotential approximation, we have exhibited the supremacy of trapping with two frequencies instead of one for the confinement of two species with widely different charge-to-mass ratios.

We have developed a computer program for simulating charged particles inside the Paul trap. Using this script, we studied the stability of particles in dependence on the trap settings, identifying whole areas of stable operating conditions. By reproducing results for trapping in one direction from reviewed scientific articles, we have confirmed the validity of our simulations. After that, we continued producing novel results by studying the stability of electrons in all spatial directions. Other than stability, we evaluated an average electron velocity throughout the trajectory dependent on the trap parameters. We have found that low electron temperatures could be achieved in the weaker electric fields. Based on these results, we propose a technique to achieve successful electron cooling by the successive change of trapping parameters. To model such a process, we must add a feature to our program representing electrons' possibility to cool themselves by interaction with ions. The program will be further developed for use in the real geometry of our trap. 
