

%%% Choose a language %%%

\newif\ifEN
\ENtrue   % uncomment this for english
%\ENfalse   % uncomment this for czech

%%% Configuration of the title page %%%

\def\ThesisTitleStyle{mff} % MFF style
%\def\ThesisTitleStyle{cuni} % uncomment for old-style with cuni.cz logo
%\def\ThesisTitleStyle{natur} % uncomment for nature faculty logo

\def\UKFaculty{Faculty of Mathematics and Physics}
%\def\UKFaculty{Faculty of Science}

\def\UKName{Charles University in Prague} % this is not used in the "mff" style

% Thesis type names, as used in several places in the title
\def\ThesisTypeTitle{\ifEN BACHELOR THESIS \else BAKALÁŘSKÁ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN MASTER THESIS \else DIPLOMOVÁ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN RIGOROUS THESIS \else RIGORÓZNÍ PRÁCE \fi}
%\def\ThesisTypeTitle{\ifEN DOCTORAL THESIS \else DISERTAČNÍ PRÁCE \fi}
\def\ThesisGenitive{\ifEN bachelor \else bakalářské \fi}
%\def\ThesisGenitive{\ifEN master \else diplomové \fi}
%\def\ThesisGenitive{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisGenitive{\ifEN doctoral \else disertační \fi}
\def\ThesisAccusative{\ifEN bachelor \else bakalářskou \fi}
%\def\ThesisAccusative{\ifEN master \else diplomovou \fi}
%\def\ThesisAccusative{\ifEN rigorous \else rigorózní \fi}
%\def\ThesisAccusative{\ifEN doctoral \else disertační \fi}



%%% Fill in your details %%%

% (Note: \xxx is a "ToDo label" which makes the unfilled visible. Remove it.)
\def\ThesisTitle{{Simulations of dynamics of ultra-cold quantum plasma}}
\def\ThesisAuthor{{Andrej Rendek}}
\def\YearSubmitted{{2022}}

% department assigned to the thesis
\def\Department{{Department of Surface and Plasma Science}}
% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{{Department}}

\def\Supervisor{{Mgr. Michal Hejduk, Ph.D.}}
\def\SupervisorsDepartment{{Department of Surface and Plasma Science}}

% Study programme and specialization
\def\StudyProgramme{{Physics}}
\def\StudyBranch{{Physics}}

\def\Dedication{%
A would like to show my appreciation for\dots}

\def\AbstractEN{%
We investigate the confinement of charged particles in an oscillating electromagnetic field, intending to trap electrons and calcium ions simultaneously. The theoretical part of this thesis focuses on the derivation of exact and effective equations of motion in a single frequency quadrupole trap. We show a general treatment of stability in such a field. Passing attained knowledge onto two-frequency trapping, which is much more suitable for confining two species with widely different charge-to-mass ratios. We follow up by studying the stability of electrons, employing computer simulations in the ideal two-frequency Paul trap. Our ambition is to identify a stable configuration minimizing electrons' temperature. We create multiple ion Coulomb crystals and examine the effect of their presence on electrons' stability. These efforts support the development of an experiment with the ambition to create and study quantum plasma. The composition of this experiment is outlined here as well.
}

\def\AbstractCS{%
}

% 3 to 5 keywords (recommended), each enclosed in curly braces.
% Keywords are useful for indexing and searching for the theses by topic.
\def\Keywords{%
{Ion trapping,} {Coulomb crystal,} {Two frequency Paul trap}
}

% If your abstracts are long and do not fit in the infopage, you can make the
% fonts a bit smaller by this setting. (Also, you should try to compress your abstract more.)
% Alternatively, consider increasing the size of the page by uncommenting the
% geometry modification in thesis.tex.
\def\InfoPageFont{}
%\def\InfoPageFont{\small}  %uncomment to decrease font size

\ifEN\relax\else
% If you are writing a czech thesis, you additionally need to fill in the
% english translation of the metadata here!
\def\ThesisTitleEN{\xxx{Thesis title in English}}
\def\DepartmentEN{\xxx{Name of the department in English}}
\def\DeptTypeEN{\xxx{Department}}
\def\SupervisorsDepartmentEN{\xxx{Superdepartment}}
\def\StudyProgrammeEN{\xxx{study programme}}
\def\StudyBranchEN{\xxx{study branch}}
\def\KeywordsEN{%
\xxx{{key} {words}}
}
\fi
